\toolnameend: A Tool for Visualizing Patterns of Spreadsheet Function Combinations

### 2016-08-10 doc/paper.tex
Fixed author block and emails

%% bare_conf.tex
%% V1.4b
%% 2015/08/26
%% by Michael Shell
%% See:
%% http://www.michaelshell.org/
%% for current contact information.
%%
%% This is a skeleton file demonstrating the use of IEEEtran.cls
%% (requires IEEEtran.cls version 1.8b or later) with an IEEE
%% conference paper.
%%
%% Support sites:
%% http://www.michaelshell.org/tex/ieeetran/
%% http://www.ctan.org/pkg/ieeetran
%% and
%% http://www.ieee.org/

@@

% *** CITATION PACKAGES ***
%
\usepackage{cite} % cite.sty was written by Donald Arseneau
% The latest version can be obtained at:
% http://www.ctan.org/pkg/cite
% The documentation is contained in the cite.sty file itself.

@@

\newcommand{\toolname}{Perquimans } \newcommand{\toolnameend}{Perquimans}
\newcommand{\toolnameposs}{Perquimans' }

@@

	% author names and affiliations
	% use a multiple column layout for up to three different
	% affiliations
	\author{ \IEEEauthorblockN{Justin A. Middleton, Emerson Murphy-Hill}
		\IEEEauthorblockA{North Carolina State University, Raleigh, USA\\ \{jamiddl2,
			ermurph3\}@ncsu.edu }}

@@

	From these stories, we can see the dire importance of understanding how people
	actually use spreadsheets. Fortunately, spreadsheet researchers have supported
	this by amassing spreadsheet collections from diverse sources, industrial to
	academic. Collections like EUSES~\cite{fisher2005euses},
	Fuse~\cite{barik2015fuse}, and the Enron corpus~\cite{hermans2015enron} have
	already fueled productive research across the field, such as in the search for
	code smells in spreadsheet
	environments~\cite{hermans2012detecting}~\cite{jansen2015code} or the creation
	of grammars to capture every possible spreadsheet program in
	Excel~\cite{aivaloglou2015grammar}. Additionally, some research has focused on
	comparing these collections, as Jansen did in 2015, to discover variation in
	spreadsheet techniques over time and location~\cite{jansen2015enron}.

@@

	This paper contributes such a tool, \toolnameend, that attempts to fulfill this
	need for better exploration by building on the aforementioned spreadsheet
	corpora. Working with spreadsheets in Excel, \toolname focuses primarily on the
	functions, like SUM and IF and hundreds more, which are the building blocks of
	formulae that determine values across a spreadsheet. The tool creates an
	interactive, exploratory visualization which captures how people combine these
	functions in a selected collection of spreadsheets, quantifying the broad
	patterns of function nesting while also being specific enough to find anomalous
	formulae and link them back to the specific cell in spreadsheets where they
	originate. \par

@@

	\section{Related Work} \label{related-work} Visualizations excel for their
	ability to communicate information quickly and
	efficiently~\cite{baeza1999modern}; the visualization of spreadsheets,
	therefore, has offered much room for research. Several previous tools work
	within a spreadsheet, seeking to clarify the connections between the visible
	values in cells, the formulae that produce them, and the dependencies between
	cells that inform them. Igarashi and colleague's fluid
	visualizations~\cite{igarashi1998fluid}, for example, do this by imposing
	lines, color, and animation over the spreadsheet to visually explain these
	tangled connections. Likewise, Clermont explored ways of grouping cells by
	color and border that were related through similar functions, neighbors, or
	references~\cite{clermont2003scalable}, and Hipfl extended this work by
	incorporating the layouts and labels of spreadsheets~\cite{hipfl2008using}.
	Other tools focus on creating visualizations external to the sheets. Hermans
	and colleagues address a spreadsheet programmer's information needs by making
	dataflow diagrams from individual sheets through the tools
	GyroSAT~\cite{hermans2011supporting} and Breviz~\cite{hermans2011breviz}.
	\toolname is among the latter class, creating standalone visualizations about
	spreadsheets instead of drawing connections inside the spreadsheet itself. In
	contrast to the other visualization work, our work focuses on representing
	functions and how they combine into complex formulae, rather than exploring the
	data structures of any particular spreadsheet.  \par

	Many other tools inspect the structure of formulae but often from the
	perspective of finding bad design.  Following Fowler's initial description of
	code smells in object-oriented design~\cite{fowler2009refactoring}, many
	researchers have applied similar observations to spreadsheets and have
	cataloged classes of problematic
	formulae~\cite{hermans2012detecting}~\cite{cunha2012towards}~\cite{asavametha2012detecting}, often accompanying them with detection tools~\cite{abreu2014smelling} and investigations into real datasets~\cite{jansen2015code}. For example, Hermans, leveraging her work with the aforementioned dataflow diagrams, has worked with assessing common inter-worksheet smells, such as feature envy and inappropriate intimacy, wherein references between worksheets (rather than cells) in a file suggest a problem~\cite{hermans2012detectinginter}. Other tools, like Badame's Excel refactoring~\cite{badame2012refactoring}, are concerned more with understanding the formula in service of changing them. However, these studies define standards of poor design and orient their searches around them. Though \toolname supports the search for bad design, as we will see, the tool itself takes a much more agnostic approach in presentation, offering examples without regard for design quality. \par

	Implicit in this study is an exploration of the Excel API and how people use
	it. Previous research, such as the first study on the Enron
	spreadsheets~\cite{hermans2015enron} or in its comparison with the EUSUS
	corpus~\cite{jansen2015enron}, quantifies how often users employ certain
	functions but not how they combine them. Researchers have applied such
	questions to other contexts as well, such as Murphy and colleague's study into
	how Java developers use Eclipse and what commands they use most
	often~\cite{murphy2006java}. Others focus on the documentation of the API
	itself: Robillard and DeLine, for example conducted a series of surveys and
	interviews to diagnose common problems with learning APIs, and they found that
	code examples, one of the focuses of this tool, is one of the most important
	aids to have in learning a new system~\cite{robillard2011field}. Though this
	paper applies the tool to the Enron corpus specifically, the tool itself is not
	limited to this; it can be readily applied to any collection of Excel
	spreadsheets a user gives it.

@@

	\subsection{Design Goals} \label{goals} In visualizing the spreadsheet data, we
	set out a few core goals for what the tool should accomplish. Furthermore, with
	these goals, we also considered the likely challenges of achieving those goals
	and particular outcomes which we considered beyond the scope of the research.
	We present both sides of these goals below:

@@

		\item [!2] \textbf{Do not overwhelm with numbers.} For popular functions
		especially, we anticipate that we'll have to show a lot of related data at
		once. Though everything should be accessible in accordance with Goal 1, we
		should design in a way that reduces noise.

@@

		\item [!3] \textbf{Do not attempt to explain formulae.} Though the tool is
		meant to foster understanding by linking pattern to example, it will not try
		to be an explanatory tool that articulates what a complex formula
		accomplishes. Rather, we will provide the structure and context, and the user
		must infer this on their own.

@@

	First, a word on the IF function. In Excel, it serves as one of the logical
	functions: given a condition as the first argument that evaluates to either
	true or false, it takes on the value of either the second or third argument,
	respectively. Or, as the documentation writes out, ``IF(Something is True, then
	do something, otherwise do something
	else)"\footnote{https://support.office.com/en-us/article/IF-function-69aed7c9-4e8a-4755-a9bc-aa8bbff73be2?ui=en-US\&rs=en-US\&ad=US}. Furthermore, the third argument is optional; if left out, the function will return a default ``FALSE" value if necessary. Examples of this function in practice can be seen in Figure~\ref{fig:ifexample}.

@@

	The visualization so far comprises two types of nodes: the circle
	\includegraphics{glossary-greenonly}, which represents a discrete function in
	the formula and is size according to its frequency relative to the number of
	times the root function appeared; and the numbered squares
	\includegraphics{glossary-blue}, which represent the positions of arguments
	within its parent function. From this, we can confirm that, of the times it was
	observed, IF can have at most three arguments passed into it, which corresponds
	with its specifications in the API. \par

	Knowing that it is the first argument which contains the conditionals, we click
	on the square labeled ``1" to explore. To save space, when the tool finds more
	than 10 unique arguments in a position, it displays only the first ten, with an
	option to display the rest by clicking the arrow
	\includegraphics{glossary-arrow}.

@@

	We see that IF contains a number of comparison operators as its first argument,
	such as = and $\le$, and boolean-returning functions, like ISNA and AND, with
	simple equality being the most common and ABS being the least. From here, we
	can further explore the common options among these functions. Clicking on the
	``=" node will yield two arguments and expanding each of them will peer into
	the range of common values of equality comparison, which Figure
	\ref{fig:fullpic} shows. \par

@@

		\item  \vspace{.25cm} \includegraphics{glossary-solidline} - \textbf{Solid
			lines} represent a function's list of parameters. As such, they connect the
		function nodes to argument nodes.

@@

		\item  \vspace{.25cm} \includegraphics[scale=.75]{glossary-lightblue} -
		\textbf{Optional argument trees} are variations of trees that show only
		formulae with a specific number of arguments. For example, the function AND
		can take any number of arguments, up to 255. By default, the tree shows data
		from AND functions with any number of arguments, but these trees, which have
		argument nodes of a lighter blue color, can show how AND was used when
		\textit{only} two arguments were passed into it, or \textit{only} four, and so
		on. See Figure~\ref{fig:optional} for how the data between tree types differs.
	\end{itemize}

@@

		\item \textit{Presentation}: The JSON files, meanwhile, feed into the
		presentation code, implemented in Javascript with much help from the
		visualization library D3\footnote{https://d3js.org/}. We chose D3 primarily
		for its accessibility: given the JSON, the visualization can be simply
		embedded into a webpage accessible through a browser. \end{itemize} A
	demonstration of \toolname can be found at
	https://github.com/DeveloperLiberationFront/Excel-Function-Visualizer.

	\subsection{Design Decisions} \label{ssec:decisions} Early in the design, we
	chose the tree form for its inherent ability to represent the parent-children
	relationship of formulae as functions and their arguments, which could be yet
	more functions. By adapting this structure for a broad range of possibilities
	for nested functions, the branching factor depends alternatively on the number
	of arguments in a function and the number of possible functions observed as an
	argument in a function. \par

	Another of our options was to represent the call trees as simple text
	hierarchies. Though this might work well for a single, uncomplicated call
	structure, we did not do this because our tool combined many of these call
	trees into a more complicated structure. To capture the different relationships
	in the tree and to quantify the frequency of arguments, we would need to be
	more verbose in our description. As such, we decided that visualizing some
	dimensions was quicker than writing them out. \par

@@

	\section{Case Study} The obvious question to pose to any visualization is this:
	What do the images tell us that the text could not? It is not enough to prove
	that something can be visualized; we must also demonstrate what we can gain
	from any particular visualization. As such, we raised a few tasks in the
	introduction in which we thought the tool could help: detecting bad smells and
	guiding spreadsheet education. To demonstrate the tool's applicability, we
	sought out places in the tree which best exemplify these tasks.

@@

	However, because \toolname makes no judgments beforehand about which designs to
	emphasize as good or bad, the tool can be used to find strange or suboptimal
	designs that exist outside these readily visible categories. Consider the
	second argument of the ``='' operator, already shown expanded in
	Figure~\ref{fig:fullpic}: among the top five argument types for the position in
	an IF function were boolean values. This corresponds to designs in which
	conditions already in a boolean form were then checked against ``=TRUE" or
	``=FALSE", such as in the formula \textit{IF(Formulas!C77=TRUE, ``Yes")}, which
	appears in one of the sheets. Though this practice does not break anything in
	itself, it nevertheless represents redundant operations in formula design and
	indicates a possible misunderstanding of how the IF function works over the
	hundreds of examples of this boolean comparison that our tool can return.

@@

	%Lookup example
	For a deeper example, the LOOKUP functions are some of the most popular in
	Excel yet come with a number of important but silent requirements for proper
	use, and, as such, they have warranted some extra attention with regards to
	spreadsheet design. In short, these functions are for ``when you need to look
	in a single row or column and find a value from the same position in a second
	row or
	column"\footnote{https://support.office.com/en-us/article/LOOKUP-function-446d94af-663b-451d-8251-369d5e3864cb}. Examples of their use can be found in Figure \ref{fig:lookupexample}, and we will cover each type in more depth to better understand the visualizations for these functions.

@@

	All of these variations may seem daunting to a beginner, but \toolname can
	help. Because of its design, users can distinguish between forms of the same
	function, as seen in Figure~\ref{fig:lookups} where one can explore both the
	frequency with which the spreadsheet programmers used these different forms --
	three-argument vector form much more than the array form -- and then retrieve
	examples of every configuration in the dataset. The same goes for the V/HLOOKUP
	functions, as shown partially in Figure~\ref{fig:vhlookups}.

@@

	Furthermore, these examples are good as well for exploring how it might bridge
	issues of bad design with education. In a recent study, Hermans and colleagues
	probed the Enron dataset for uses of VLOOKUP and HLOOKUP
	specifically~\cite{hermans2015detecting}. The issue at hand was the optional
	fourth argument, which requires that rows are sorted lest it return inaccurate
	results. For someone seeking to increase awareness of this problem, it first
	allows them to answer the question of significance -- ``How often do
	spreadsheet programmers use this function, with these particular parameters?"
	Though \toolname does not discern whether neighboring columns are sorted, it
	can nevertheless guide the educator to several examples in production where
	this problem surfaces. \par

@@

	The participants were most vocal about their findings when they discovered, by
	purpose or accident, instances of heinous design in the dataset. Many times,
	they noted these examples shortly after opening a new tree, suggesting that the
	distinctive visual signatures as discussed in Section \ref{badsmells} was a
	salient force in their exploration. For example, when P3 encountered the
	initial design of the SUM tree, as seen in Figure \ref{fig:sum}, they readily
	diagnosed that either a spreadsheet author was ``totally incompetent, or [the
	formula was] just something that's weird." Even though the visualization
	struggles to scale well for this formula, it calls clearly to a problem in the
	underlying spreadsheets -- awkward formulae make awkward trees. Other than
	these visual cues, value errors, prefixed with \#s, also appear in the tree,
	which P4 said could be useful for tracking these problems if marked well.

	\begin{wrapfigure}{L}{.25\textwidth} \centering
		\includegraphics[width=.20\textwidth]{SUM}  \caption{An abundance of parameters
			in SUM, which was a commonly noted smell in the user study} \label{fig:sum}
	\end{wrapfigure}

	Additionally, because the Enron files had names with the author as a prefix, P2
	commented that companies could use the tool to follow up with repeat offenders.
	For instance, they first encountered a particular author's name when
	investigating an AVERAGE function with 18 distinct arguments for which our
	participant could not discern any logical order even after opening the source
	spreadsheet; the same name popped up again as P2 looked at a RATE function
	relying on several hard-coded values. P2 said that, if they were a manager or
	project leader, this recurrence was a sign to talk to the author personally and
	figure out why these problems were there, as to prevent more in the future.
	They also added that these problems would grow worse if this particular author
	left the company and no one remained who understood these formulae, a concern
	that aligns with Hermans' transfer scenarios \cite{hermans2011supporting}.

	The detection of bad design, however, is certainly not limited to a
	visualization tool. Some users remarked that a tool which printed a list of
	errors could work just as well for that purpose, and smell checkers can and
	have been directly built into spreadsheets to point out bad design. However,
	users also said that this visual interface is better when searching for new
	smells. Because the exploratory design makes no qualitative judgment and
	visualizes benign designs with the problematic, it allows for flexible
	interpretation of what a smell is; that is, a user might not realize a certain
	design is poor until they see it in the tree and, at the point, find specific
	examples of where it occurs in the spreadsheets.

	\subsection{Use in Education} Users found \toolname's inclusion of examples to
	be especially useful since none of the them considered themselves to be experts
	of Excel, their self-reported proficiencies ranging from ``fairly familiar"
	(P2) to ``not extremely familiar" (P4). For example, as P1 explored the VLOOKUP
	function, they said they could use the tool to collect more examples than the
	documentation offers even though they knew the documentation more verbosely
	explained its function signature. Furthermore, in the process of exploring
	trees, P1 and P3 also gravitated towards some functions they had not used
	before, such as PV and IPMT, looking for new features in Excel to explore and
	learn. \par

@@

	\subsection{Study Limitations} \label{subsec:studylimitations} Though we did not start the study by telling
	that the data source came from Enron, two users studied the file names to
	deduce its roots in finance, if not Enron itself. Because of these notorious
	connotations, then, they might have been inclined from this information to
	orient their exploration around finance-related functions, potentially limiting
	their findings or perceived users for the tool. However, one participant also
	said that, without this essential context, they could not have intelligently
	explored a dataset in the first place. Comments like these point out the
	trade-offs of unguided exploration, too: though it might not restrict them to a
	single purpose, it might, at the same time, not afford them any mindset at all
	with which to interpret the data. \par

@@

	A threat to our goals arises from the problem of too much data,
	particularly in the popular functions, like SUM and IF. We realized early that
	these functions would overwhelm the user with options in large datasets, so we
	included features, such as the expansion arrows and the use of generic types
	instead of values, that would manage the variation. Nevertheless, P3 remarked
	that they specifically avoided the most popular functions, anticipating a flood
	of nodes from which they could salvage nothing interesting. P2 corroborated
	this by explaining how they gravitated toward exploring specific examples in
	particular value nodes over navigating the full, crowded tree, but they also
	said that the tool at its best when it compares either every possibility of an
	argument side-by-side or nothing on that branch at all. Despite our
	precautions, balancing the desire to show a lot with the desire not to
	overwhelm remains a challenge. \par

	Some visual cues turned out to be more confusing than helpful, such as
	the index number on each argument node. As discussed in Section~\ref{subsec:studylimitations},
	users often mistook it for other features, like total number of arguments in the
	function. Though we kept this feature present for all studies, future versions
	of the tool would refine features like this to reduce confusion.

	\subsection{Informative Qualification}

>>>>

%% bare_conf.tex
%% V1.4b
%% 2015/08/26
%% by Michael Shell
%% See:f
%% http://www.michaelshell.org/
%% for current contact information.
%%
%% This is a skeleton file demonstrating the use of IEEEtran.cls
%% (requires IEEEtran.cls version 1.8b or later) with an IEEE
%% conference paper.
%%
%% Support sites:
%% http://www.michaelshell.org/tex/ieeetran/
%% http://www.ctan.org/pkg/ieeetran
%% and
%% http://www.ieee.org/

@@

% *** CITATION PACKAGES ***
%
\usepackage{cite} % cite.sty was written by Donald Arseneau
% The latest version can be obtained at:
% http://www.ctan.org/pkg/cite
% The documentation is contained in the cite.sty file itself.
\usepackage{xspace}

@@

\newcommand{\toolname}{Perquimans\xspace} \newcommand{\toolnameend}{Perquimans}
\newcommand{\toolnameposs}{Perquimans' }

@@

	% author names and affiliations
	% use a multiple column layout for up to three different
	% affiliations
	\author{ \IEEEauthorblockN{Justin A. Middleton and Emerson Murphy-Hill}
		\IEEEauthorblockA{North Carolina State University, Raleigh, USA\\ jamiddl2@ncsu.edu, emerson@csc.ncsu.edu }}

@@

	From these stories, we can see the dire importance of understanding how people
	actually use spreadsheets. Fortunately, spreadsheet researchers have supported
	this by amassing spreadsheet collections from diverse sources, industrial to
	academic. Collections like EUSES~\cite{fisher2005euses},
	Fuse~\cite{barik2015fuse}, and the Enron corpus~\cite{hermans2015enron} have
	already fueled productive research across the field, such as in
	Hermans'~\cite{hermans2012detecting} and Jansen's~\cite{jansen2015code} search
	for code smells in spreadsheet environments or Aivaloglou and colleague's work
	on making a grammar to capture every possible spreadsheet program in
	Excel~\cite{aivaloglou2015grammar}. Additionally, some research has focused on
	comparing these collections, as Jansen did in 2015, to discover variation in
	spreadsheet techniques over time and location~\cite{jansen2015enron}.

@@

	This paper contributes such a tool, \toolname, that attempts to fulfill this
	need for better exploration by building on the aforementioned spreadsheet
	corpora. Working with spreadsheets in Excel, \toolname focuses primarily on the
	functions, like SUM and IF and hundreds more, which are the building blocks of
	formula and programs that determine values across a spreadsheet. The tool
	creates an interactive, exploratory visualization which captures how people
	combine these functions in a selected collection of spreadsheets, quantifying
	the broad patterns of function nesting while also being specific enough to find
	anomalous formulae and link them back to the specific cell in spreadsheets
	where they originate. \par

@@

	\section{Related Work} \label{related-work} Visualizations excel for their
	ability to communicate information quickly and
	efficiently~\cite{baeza1999modern}; the visualization of spreadsheets,
	therefore, has offered much room for research. Several previous tools work
	within a spreadsheet, seeking to clarify the connections between the visible
	values in cells, the formulae that produce them, and the dependencies between
	cells that inform them. Igarashi and colleague's fluid
	visualizations~\cite{igarashi1998fluid}, for example, do this by imposing
	lines, color, and animation over the spreadsheet to visually explain these
	tangled connections. Likewise, Clermont explored ways of grouping cells by
	color and border that were related through similar functions, neighbors, or
	references~\cite{clermont2003scalable}, and Hipfl extended this work by
	incorporating the layouts and labels of spreadsheets~\cite{hipfl2008using}.
	Other tools focus on creating visualizations external to the sheets. Hermans
	and colleagues address a spreadsheet programmer's information needs by making
	dataflow diagrams from individual sheets through the tools
	GyroSAT~\cite{hermans2011supporting} and Breviz~\cite{hermans2011breviz}.
	\toolname is among the latter class, creating standalone visualizations about
	spreadsheets instead of drawing connections inside the spreadsheet itself. In
	contrast to Hermans' work, our work focuses on representing functions and how
	they comprise complex formula, rather than exploring the data structures of any
	particular spreadsheet.  \par

	Many other tools inspect the structure of formulae but often from the
	perspective of finding bad design.  Following Fowler's initial description of
	code smells in object-oriented design~\cite{fowler2009refactoring}, many
	researchers have applied similar observations to spreadsheets and have
	cataloged classes of problematic
	formulae~\cite{hermans2012detecting}~\cite{cunha2012towards}~\cite{asavametha2012detecting}, often accompanying them with detection tools~\cite{abreu2014smelling} and investigations into real datasets~\cite{jansen2015code}. For example, Hermans, leveraging her work with the aforementioned dataflow diagrams, has worked with assessing common inter-worksheet smells, such as feature envy and inappropriate intimacy, wherein references between worksheets (rather than cells) in a file suggest a problem~\cite{hermans2012detectinginter}. Other tools, like Badam's Excel refactoring~\cite{badame2012refactoring}, are concerned more with understanding the formula in service of changing them. However, these studies define standards of poor design and orient their searches around them. Though \toolname supports the search for bad design, as we will see, the tool itself takes a much more agnostic approach in presentation, offering examples without regard for design quality. \par

	Implicit in this study is an exploration of the Excel API and how people use
	it. Previous research, such as the first study on the Enron
	spreadsheets~\cite{hermans2015enron} or in its comparison with the EUSUS
	corpus~\cite{jansen2015enron}, quantify how ofter users employ certain
	functions but not how they combine them. Researchers have applied such
	questions to other contexts as well, such as Murphy and colleague's study into
	how Java developers use Eclipse and what commands they use most
	often~\cite{murphy2006java}. Others focus on the documentation of the API
	itself: Robillard and DeLine, for example conducted a series of surveys and
	interviews to diagnose common problems with learning APIs, and they found that
	code examples, one of the focuses of this tool, is one of the most important
	aids to have in learning a new system~\cite{robillard2011field}. Though this
	paper applies the tool to the Enron corpus specifically, the tool itself is not
	limited to this; it can be readily applied to any collection of Excel
	spreadsheets a user gives it.

@@

	\subsection{Design Goals} \label{goals} In visualizing the spreadsheet data, we
	follow a few core goals for what the tool should accomplish. Furthermore, with
	these goals, we also considered the likely challenges of achieving those goals
	and particular outcomes which we considered beyond the scope of the research.
	We present both sides of these goals below:

@@

		\item [!2] \textbf{Do not overwhelm with numbers.} For popular functions
		especially, we anticipated that we'd have to show a lot of related data at
		once. Though everything should be accessible in accordance with Goal 1, we
		should design in a way that reduces noise.

@@

		\item [!3] \textbf{Do not attempt to explain formulae.} Though the tool will
		to foster understanding by linking pattern to example, it will not try to be
		an explanatory tool that articulates what a complex formula accomplishes. Rather,
		we will provide the structure and context, and the user	must infer this on their own.

@@

	First, a word on the IF function. In Excel, it serves as one of the logical
	functions: given a condition as the first argument that evaluates to either
	true or false, it takes on the value of either the second or third argument,
	respectively. Or, as the documentation writes out, "IF(Something is True, then
	do something, otherwise do something
	else)"\footnote{https://support.office.com/en-us/article/IF-function-69aed7c9-4e8a-4755-a9bc-aa8bbff73be2?ui=en-US\&rs=en-US\&ad=US}. Furthermore, the third argument is optional; if left out, the function will return a default "FALSE" value if necessary. Examples of this function in practice can be seen in Figure~\ref{fig:ifexample}.

@@

	The visualization so far comprises two types of nodes: the circle
	\includegraphics{glossary-greenonly}, which represents a discrete function in
	the formula and is size according to its frequency relative to the number of
	times to root function appeared; and the numbered squares
	\includegraphics{glossary-blue}, which represent the positions of arguments
	within its parent function. From this, we can confirm that, of the times it was
	observed, IF can have at most three arguments passed into it, which corresponds
	with its specifications in the
	API\footnote{https://support.office.com/en-us/article/IF-function-69aed7c9-4e8a-4755-a9bc-aa8bbff73be2}. \par

	Knowing that it is the first argument which contains the conditionals, we click
	on the square labeled ``1" to explore. To save space, when more than 10 unique
	arguments were found in a position, the tool displays only the first
	ten, with an option to display the rest by clicking the arrow
	\includegraphics{glossary-arrow}.

@@

	We see that IF contains a number of comparison operators as its first argument,
	such as = and $\le$, and boolean-returning functions, like ISNA and AND, with
	simple equality being the most common and some use as ABS being the least seen
	of everything actually used. From here, we can further explore the common options among these functions.
	Clicking on the ``=" node will yield two arguments and expanding each of them
	will peer into the range of common values of equality comparison, which Figure
	\ref{fig:fullpic} shows. \par

@@

		\item  \vspace{.25cm} \includegraphics{glossary-solidline} - \textbf{Solid
			lines} represent a function's acceptance of its parameters. As such, they
		connect the function nodes to argument nodes.

@@

		\item  \vspace{.25cm} \includegraphics[scale=.75]{glossary-lightblue} -
		\textbf{Optional argument trees} are variations of trees that show only
		formulae with a specific number of arguments. For example, the function AND
		can take any number of arguments, up to 255. By default, the tree shows data
		from AND functions with any number of argument, but these trees, marked with a
		darker blue color, can show how AND was used when \textit{only} two arguments
		were passed into it, or \textit{only} four, and so on. See
		Figure~\ref{fig:optional} for how the data between tree types differs.
	\end{itemize}

@@

		\item \textit{Presentation}: The JSON files, meanwhile, feed into the
		presentation code, implemented in Javascript with much help from the
		visualization library D3\footnote{https://d3js.org/}. We chose D3 primarily
		for its accessibility: given the JSON, the visualization can be simply
		embedded into a webpage accessible through a browser. \end{itemize} The code and a
	demonstration of \toolname can be found at
	https://github.com/DeveloperLiberationFront/Excel-Function-Visualizer.

	\subsection{Design Decisions} \label{ssec:decisions} Early in the design, we
	chose the tree form for its inherent ability to represent the parent-children
	relationship of formulae as functions and their arguments, which could be yet
	more functions. By adapting this structure to accommodate a broad range of
	possibilities for nested functions, the branching factor depends alternatively
	on the number of arguments in a function and the number of possible functions
	observed as an argument in a function. \par

@@

	\section{Case Study} The obvious question to pose to any visualization is this:
	What do the images tell us that the text could not? It is not enough to prove
	that something can be visualized; we must also demonstrate what we can be
	gained from any particular visualization. As such, we raised a few tasks in the
	introduction in which we thought the tool could help: detecting bad smells and
	guiding spreadsheet education. To demonstrate the tool's applicability, we
	sought out places in the tree which best exemplify these tasks.

@@

	However, because \toolname makes no judgments beforehand about which designs to
	emphasize as good or bad, the tool can be used to find strange or suboptimal
	designs that exist outside these obvious categories. Consider the second
	argument of the ``='' operator, already shown expanded in
	Figure~\ref{fig:fullpic}: among the top five argument types for the position in
	an IF function were boolean values. This corresponds to designs in which
	conditions already in a boolean form were then checked against "=TRUE" or
	"=FALSE", such as in the formula \textit{IF(Formulas!C77=TRUE, "Yes")}, which
	appears in one of the sheets. Though this practice does not break anything in
	itself, it nevertheless represents redundant operations in formula design and
	indicates a possible misunderstanding of how the IF function works over the
	hundreds of examples of this boolean comparison that our tool can return.

@@

	%Lookup example
	For a deeper example, the LOOKUP functions are some of the most popular in
	Excel yet come with a number of important but silent requirements for proper
	use, and, as such, they have warranted some extra attention with regards to
	spreadsheet design. In short, these functions are for "when you need to look in
	a single row or column and find a value from the same position in a second row
	or column"\footnote{https://support.office.com/en-us/article/LOOKUP-function-446d94af-663b-451d-8251-369d5e3864cb}. Examples of their use can be found in Figure \ref{fig:lookupexample}, and we will cover each type
	in more depth to better understand the visualizations for these functions.

@@

	All of these variations may seem daunting to a beginner, but \toolname can
	help. Because of its design, users can distinguish between forms of the same
	function, as seen in Figure~\ref{fig:lookups} where the one can explore both
	the frequency with which the spreadsheet programmers used these different forms
	-- three-argument vector form much more than the array form -- and then
	retrieve examples of every configuration in the dataset. The same goes for the
	V/HLOOKUP functions, as shown partially in Figure~\ref{fig:vhlookups}.

@@

	Furthermore, these examples are good as well for exploring how it might bridge
	issues of bad design with education. In a recent study, Hermans, Aivaloglou,
	and Jansen probed the Enron dataset for uses of VLOOKUP and HLOOKUP
	specifically. The issue at hand was the optional fourth argument, which
	requires that rows are sorted lest it return inaccurate results
	\cite{hermans2015detecting}. For someone seeking to increase awareness of this
	problem, it first allows them to answer the question of significance -- "How
	often do spreadsheet programmers use this function, with these particular
	parameters?" Though \toolname does not discern whether neighboring columns are
	sorted, it can nevertheless guide the educator to several examples in
	production where this problem surfaces. \par

@@

	The participants were most vocal about their findings when they discovered, by
	purpose or accident, instances of heinous design in the dataset. Many times,
	they noted these examples shortly after opening a new tree, suggesting that the
	distinctive visual signatures as discussed in Section \ref{badsmells} was a
	salient force in their exploration. For example, when P3 encountered the
	initial design of the SUM tree, as seen in Figure \ref{fig:sum}, they readily
	diagnosed that either a spreadsheet author was "totally incompetent, or [the
	formula was] just something that's weird." Other than these visual cues, value
	errors, prefixed with \#s, also appear in the tree, which P4 said could be
	useful for tracking these problems if marked well.

	\begin{wrapfigure}{O}{.25\textwidth}
		\centering \includegraphics[width=.20\textwidth]{SUM}  \caption{An abundance of parameters in SUM.
			Encountering this, P4 remarked, "Woah."} \label{fig:sum} \end{wrapfigure}

	Additionally, because the files in the Enron set had name with the author as a
	prefix, P2 commented that they could use the tool in a company setting to
	follow up with repeat offenders. For instance, they first encountered a
	particular author's name when investigating an AVERAGE function with 18
	distinct arguments for which our participant could not discern any logical
	order, even after opening the source spreadsheet; the same name popped up again
	as P2 looked at a RATE function relying on several hard-coded values. P2 said
	that, if they were a manager or project leader, this recurrence was a sign to
	talk to the author personally and figure out why these problems were there, as
	to prevent more in the future. They also added that these problems would grow
	worse if this particular author left the company and no one remained who
	understood these formulae, a concern that aligns with Hermans' transfer
	scenarios \cite{hermans2011supporting}.

	The detection of bad design, however, is certainly not limited to a
	visualization tool, and some users remarked that a tool which printed a list of
	errors could suffice just as well for that purpose. However, they also remarked
	that this visual interface better supports the case when searching for new and
	previously unknown smells. Because the exploratory design makes no
	qualitative judgment and visualizes benign designs with the problematic, it
	therefore allows for flexible interpretation of what is a smell; that is, a
	user might not realize a certain design is suboptimal until they see it in the
	tree and, at the point, find specific examples of where it occurs in the
	spreadsheets.

	\subsection{Use in Education} Users found \toolname's inclusion of examples to
	be especially useful since none of the them considered themselves to be experts
	of Excel, their self-reported proficiencies ranging from "fairly familiar" (P2)
	to "not extremely familiar" (P4). For example, as P1 explored the VLOOKUP
	function, they said they could use the tool to collect more examples than the
	documentation offers even though they knew the documentation more verbosely
	explained its function signature. Furthermore, in the process of exploring
	trees, P1 and P3 also gravitated towards some functions they had not used
	before, such as PV and IPMT, looking for new features in Excel to explore and
	learn. \par

@@

	\subsection{Study Limitations} Though we did not start the study by telling
	that the data source came from Enron, two users studied the file names to
	deduce its roots in finance, if not Enron itself. Because of these notorious
	connotations, then, they might have been inclined from this information to
	orient their exploration around finance-related functions, potentially limiting
	their findings or perceived users for the tool. However, one participant also said that,
	without this essential context, they could not have intelligently explored a dataset
	in the first place. Comments like these point out the trade-offs of unguided
	exploration, too: though it might not restrict them to a single purpose, it
	might, at the same time, not afford them any mindset at all with which to
	interpret the data. \par

@@

	We realized early that some of the popular functions would have thousands of variations in large datasets, and showing all of
	these variations at once would hopelessly overwhelm the user. We therefore made
	a number of choices, such as the expansion arrows and the use of generic types
	in place of specific primitive values, that would reduce the variation to
	something more manageable.

	Nevertheless, during the user study, P3 remarked that they specifically avoided
	the most popular functions, anticipating a flood of nodes from which they could
	salvage nothing interesting. P2 corroborated this by explaining how they
	gravitated toward exploring specific examples in particular value nodes over
	navigating the full, crowded tree, but they complicate it further by also
	saying that the tool at its best when it shows either every possibility of an
	argument side-by-side (for comparison) or nothing on that branch at all, a
	balance that is hard to strike with node-dense trees.

	A threat to our goals, then, arises from the problem of too much data,
	particularly in the popular functions, like SUM and IF. Despite our
	precautions, balancing the desire to show every single possibility of design
	with the desire not to overwhelm remains a challenge. \par

	\subsection{Informative Qualification} As we found in the case study, the tool
	excels at making formula examples available throughout a tree. 	
	However, the benefits described in the previous sections imply a hidden
	complication: if the tool educates through example but does not explicitly
	distinguish between good design and bad design, how can we be sure that people
	will not inadvertently learn bad design from it? In the user study, we
	discussed its use in education and smell detection by bringing up the issues
	surrounding the LOOKUP functions; this discussion, though, presupposes a
	distinguishing eye that knew which examples practiced good design and which
	exhibited the common errors. A beginner might not know the difference.
	Robillard and Deline, in their study, remarked that good API examples should
	demonstrate best practices, but, right now, we have no way to assure this with
	\toolname. As with the problems in achieving unbiased exploration, correcting
	for this danger may entail a reevaluation of design goals.

	\section{Future Work} For these user and case studies, we relied solely on
	Enron's spreadsheets. However, Enron's focus on finances and energy represents
	only one context in which people use spreadsheets. Other corpora, like EUSES
	and Fuse, draw from other domains and industries, and, as Jansen found in his
	comparison of spreadsheets Enron and EUSES \cite{jansen2015enron}, different
	companies rely on different functions. Though the use of a single corpora
	nevertheless suffices for evaluation, we may yet be able to find more
	interesting applications and uses for the tool in other types of spreadsheet.
	\par
